How cryptic is cryptic diversity? Machine learning approaches to classifying morphological variation in the Pacific Pond Turtle (\textit{Emys marmorata})

### 2016-01-20 doc/draft_turtle.tex
new data group. moved clusters out. redoing work

@@

\subsection{Unsupervised learning}
Comparison of gap statistic values from PAM clustering show that the optimal, minimal number of clusters is most likely one (Fig. \ref{fig:gap}). There is some ambiguity in choice because, although it is not statistically different from a solution with only one group, the solution with two groupings does have the greatest mean gap statistic. However, there is no geographical signal in the results of this clustering solution (Fig. \ref{fig:gap_map}). Because of this, we assert that there is no means of naturally partitioning plastron shape into distinct subgroups with out reference to external information.
% PAM stuff
% figure

@@

@@

@@

% generalization
Mean AUC values for the model generalizations, in most cases, are approximately equal to the observed AUC values from the training data set (Table \ref{tab:comp}). The  cases in which the AUC from the  generalizations is less than the observed, indicate poor model fit and a poor classification scheme. AUC values from model generalization, or estimating testing data set membership, does not indicate a clear ``best'' classification scheme (Fig. \ref{fig:gen_hist}). Although the scheme with two species has the greatest AUC point estimate for each modeling approach, this scheme is not significantly greater than any other except in some limited cases (e.g. LDA, Table \ref{tab:gen_tests}). 
% misfits
Differences in mean shape between correctly and incorrectly classified observations from test set frequently were statistically significant, though there are exceptions. Again, this test was to determine if the mean shapes were statistically different or not. The frequency of these results, however, is important because it means that the different models are poor predictors of class membership. This may be because differences in plastron shape do not align with the any of the hypothesized classification schemes.

@@

@@

\clearpage

\begin{table}[ht]
  \centering
  \caption{Results of comparisons between correctly and incorrectly classified observations from the testing data set, defined P(distance - simulated \(>\) 0). For each scheme, the classifications with at least 10 observations were tested. This was done for each of the three modeling techniques included in this study: random forest (RF), multinomial logistic regression (MLR), and linear discriminate analysis (LDA).}
  \begin{tabular}{ l | l r r r }
    \hline
    Scheme & Class & RF & MLR & LDA \\
    \hline
    \hline
    \multirow{2}{*}{Morph 1} & CCR & 0.77 & 1.00 & 1.00 \\ 
    & marm & 0.87 & 0.93 & 1.00 \\ 
    \hline
    \multirow{2}{*}{Morph 2} & CCR & 0.88 & 1.00 & 1.00 \\ 
    & marm & 1.00 & 1.00 & 0.98 \\ 
    \hline
    \multirow{2}{*}{Mito 1} & CCR & 0.94 & 0.99 & 0.98 \\ 
    & marm & 0.99 & 0.98 & 0.47 \\ 
    \hline
    \multirow{2}{*}{Mito 2} & marm & 0.85 & 0.97 & 0.96 \\ 
    & pall & 0.94 & 1.00 & 1.00 \\ 
    \hline
    \multirow{2}{*}{Mito 3} & 1 & 0.40 & 0.92 & 0.90 \\ 
    & 3 & 0.97 & 0.99 & 1.00 \\ 
    \hline
    \multirow{2}{*}{Nuclear} & marm & 1.00 & 1.00 & 1.00 \\ 
    & pall & 0.99 & 1.00 & 1.00 \\ 
    \hline
  \end{tabular}
  \label{tab:miss_tests}
\end{table}

\clearpage

\begin{figure}[ht]
  \centering
  \includegraphics[height = \textheight, width = \textwidth, keepaspectratio = true]{figure/plastra}
  \caption{Depiction of general plastral shape of \textit{E. marmorata} and position of the 19 landmark used in this study. Anterior is towards the top of the figure.}
  \label{fig:plastra}
\end{figure}

\begin{figure}[ht]
  \centering
  \includegraphics[height = \textheight, width = \textwidth, keepaspectratio = true]{figure/gap_res}
  \caption{Results from PAM clustering of the Riemannian shape distance for 8 different number of clusters. Vertical lines are 1 standard deviation of the mean determined from 500 resamples.}
  \label{fig:gap}
\end{figure}

\begin{figure}[ht]
  \centering
  \includegraphics[height = \textheight, width = \textwidth, keepaspectratio = true]{figure/gap_map}
  \caption{Comparison of geographic distribution of clustered observations from the 2 clustering PAM solution. Colour and shape correspond to each of the groups. There is clearly no geographic signal in the data.}
  \label{fig:gap_map}
\end{figure}

\begin{figure}[ht]
  \centering
  \includegraphics[height = \textheight, width = \textwidth, keepaspectratio = true]{figure/sel_val}
  \caption{Graphical representation of the AUC values from model selection for multinomial logistic regression and linear discriminate analysis, respecitively. AUC model selection is based on greatest AUC value. The horizontal axis corresponds to the cummulative number of axes included in the model of interest. A highlighted point corresponds to the AUC best model for that classification scheme.}
  \label{fig:sel}
\end{figure}

\begin{figure}[ht]
  \centering
  \includegraphics[height = \textheight, width = \textwidth, keepaspectratio = true]{figure/var_imp}
  \caption{Variable importance from the random forest models for each of the six classification schemes. Importance is measured as the mean decrease in Gini Index, which is a measure of the strength by which that variable determines CART structure. Indices that are farther to the right indicate greater variable importance.}
  \label{fig:var_imp}
\end{figure}

\begin{figure}[ht]
  \centering
  \includegraphics[height = \textheight, width = \textwidth, keepaspectratio = true]{figure/gen_res}
  \caption{Bootstrap distributions for genearlized AUC values for each of the classification schemes. Each row corresponds to a different modeling approach: multinomial logistic regression (MLR), random forest (RF), linear discriminate analysis (LDA), and LDA using best variables from random forest. Each distribution corresponds to 1000 bootstrap replicates.}
  \label{fig:gen_hist}
\end{figure}

\begin{figure}[ht]
  \centering
  \includegraphics[height = \textheight, width = \textwidth, keepaspectratio = true]{figure/seven_boot}
  \caption{Distribution of 1000 bootstrap replicates of the out-of-sample AUC values from the secondary, multi-species dataset. The AUC values are from the random forest (RF), linear discriminate analysis (LDA), and multinomial logistic regression (MLR). AUC values range between 0.5 (no better than random) and 1 (perfect classification).}
  \label{fig:seven_boot}
\end{figure}

>>>>

\begin{figure}[ht]
  \centering
  \includegraphics[height = 0.5\textheight, width = \textwidth, keepaspectratio = true]{figure/plastra}
  \caption{Depiction of general plastral shape of \textit{E. marmorata} and position of the 19 landmark used in this study. Anterior is towards the top of the figure.}
  \label{fig:plastra}
\end{figure}

@@

\subsection{Unsupervised learning}

\begin{figure}[ht]
  \centering
  \includegraphics[height = 0.5\textheight, width = \textwidth, keepaspectratio = true]{figure/gap_res}
  \caption{Results from PAM clustering of the Riemannian shape distance for 8 different number of clusters. Vertical lines are 1 standard deviation of the mean determined from 500 resamples.}
  \label{fig:gap}
\end{figure}

Comparison of gap statistic values from PAM clustering show that the optimal, minimal number of clusters is most likely one (Fig. \ref{fig:gap}). There is some ambiguity in choice because, although it is not statistically different from a solution with only one group, the solution with two groupings does have the greatest mean gap statistic. However, there is no geographical signal in the results of this clustering solution (Fig. \ref{fig:gap_map}). Because of this, we assert that there is no means of naturally partitioning plastron shape into distinct subgroups with out reference to external information.

\begin{figure}[ht]
  \centering
  \includegraphics[height = 0.5\textheight, width = \textwidth, keepaspectratio = true]{figure/gap_map}
  \caption{Comparison of geographic distribution of clustered observations from the 2 clustering PAM solution. Colour and shape correspond to each of the groups. There is clearly no geographic signal in the data.}
  \label{fig:gap_map}
\end{figure}

@@

\begin{figure}[ht]
  \centering
  \includegraphics[height = 0.5\textheight, width = \textwidth, keepaspectratio = true]{figure/var_imp}
  \caption{Variable importance from the random forest models for each of the six classification schemes. Importance is measured as the mean decrease in Gini Index, which is a measure of the strength by which that variable determines CART structure. Indices that are farther to the right indicate greater variable importance.}
  \label{fig:var_imp}
\end{figure}

@@

\begin{figure}[ht]
  \centering
  \includegraphics[height = 0.5\textheight, width = \textwidth, keepaspectratio = true]{figure/gen_res}
  \caption{Bootstrap distributions for genearlized AUC values for each of the classification schemes. Each row corresponds to a different modeling approach: multinomial logistic regression (MLR), random forest (RF), linear discriminate analysis (LDA), and LDA using best variables from random forest. Each distribution corresponds to 1000 bootstrap replicates.}
  \label{fig:gen_hist}
\end{figure}

@@

\begin{figure}[ht]
  \centering
  \includegraphics[height = 0.5\textheight, width = \textwidth, keepaspectratio = true]{figure/sel_val}
  \caption{Graphical representation of the AUC values from model selection for multinomial logistic regression and linear discriminate analysis, respecitively. AUC model selection is based on greatest AUC value. The horizontal axis corresponds to the cummulative number of axes included in the model of interest. A highlighted point corresponds to the AUC best model for that classification scheme.}
  \label{fig:sel}
\end{figure}

Mean AUC values for the model generalizations, in most cases, are approximately equal to the observed AUC values from the training data set (Table \ref{tab:comp}). The  cases in which the AUC from the  generalizations is less than the observed, indicate poor model fit and a poor classification scheme. AUC values from model generalization, or estimating testing data set membership, does not indicate a clear ``best'' classification scheme (Fig. \ref{fig:gen_hist}). Although the scheme with two species has the greatest AUC point estimate for each modeling approach, this scheme is not significantly greater than any other except in some limited cases (e.g. LDA, Table \ref{tab:gen_tests}).

Differences in mean shape between correctly and incorrectly classified observations from test set frequently were statistically significant, though there are exceptions. Again, this test was to determine if the mean shapes were statistically different or not. The frequency of these results, however, is important because it means that the different models are poor predictors of class membership. This may be because differences in plastron shape do not align with the any of the hypothesized classification schemes.

@@

\begin{figure}[ht]
  \centering
  \includegraphics[height = 0.5\textheight, width = \textwidth, keepaspectratio = true]{figure/seven_boot}
  \caption{Distribution of 1000 bootstrap replicates of the out-of-sample AUC values from the secondary, multi-species dataset. The AUC values are from the random forest (RF), linear discriminate analysis (LDA), and multinomial logistic regression (MLR). AUC values range between 0.5 (no better than random) and 1 (perfect classification).}
  \label{fig:seven_boot}
\end{figure}

@@

%\begin{table}[ht]
%  \centering
%  \caption{Results of comparisons between correctly and incorrectly classified observations from the testing data set, defined P(distance - simulated \(>\) 0). For each scheme, the classifications with at least 10 observations were tested. This was done for each of the three modeling techniques included in this study: random forest (RF), multinomial logistic regression (MLR), and linear discriminate analysis (LDA).}
%  \begin{tabular}{ l | l r r r }
%    \hline
%    Scheme & Class & RF & MLR & LDA \\
%    \hline
%    \hline
%    \multirow{2}{*}{Morph 1} & CCR & 0.77 & 1.00 & 1.00 \\ 
%    & marm & 0.87 & 0.93 & 1.00 \\ 
%    \hline
%    \multirow{2}{*}{Morph 2} & CCR & 0.88 & 1.00 & 1.00 \\ 
%    & marm & 1.00 & 1.00 & 0.98 \\ 
%    \hline
%    \multirow{2}{*}{Mito 1} & CCR & 0.94 & 0.99 & 0.98 \\ 
%    & marm & 0.99 & 0.98 & 0.47 \\ 
%    \hline
%    \multirow{2}{*}{Mito 2} & marm & 0.85 & 0.97 & 0.96 \\ 
%    & pall & 0.94 & 1.00 & 1.00 \\ 
%    \hline
%    \multirow{2}{*}{Mito 3} & 1 & 0.40 & 0.92 & 0.90 \\ 
%    & 3 & 0.97 & 0.99 & 1.00 \\ 
%    \hline
%    \multirow{2}{*}{Nuclear} & marm & 1.00 & 1.00 & 1.00 \\ 
%    & pall & 0.99 & 1.00 & 1.00 \\ 
%    \hline
%  \end{tabular}
%  \label{tab:miss_tests}
%\end{table}
